A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-07-27 survey.tex
Added a new (tentative) version of the complexity of StarCraft

Another way  to measure the complexity  of the game is  by looking at
the branching factor, $b$, and the depth of the game, $d$, as proposed in~\cite{Gaby}, with a total game complexity of $b^d$. In Chess, $b \approx 35$ and $d \approx 80$. In more complex games, like Go, $b \approx 30 - 300$, and $d \approx 150 - 200$. In order to determine the branching factor in StarCraft when an AI plays it, we must have in mind, that the AI can issue actions simultaneously to as many units in the game as desired. Thus, considering that, in a typical game, a player controls between 50 to 200 units, the branching factor would be between $u^{50}$ and $u^{200}$, where $u$ is the number of actions each unit can execute. Assuming a conservative value for $u$ of about $30$ (moving to nearby positions, attacking units in range, using special abilities, etc.), we obtain a branching factor of $b$ between  $30^{50}$ and $30^{200}$. Now, to compute $d$, we simply consider the fact that typical games last for about 25 minutes, which results in $d \approx 36000$ (25 minutes $\times$ 60 seconds $\times$ 24 iterations per second).

@@

% CBR:
Also falling into the machine-learning category, a significant group of researchers has explored case-based reasoning (CBR) \cite{Aamodt94CBR} approaches for strategic decision making. For example Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval in the Wargus domain. Hsieh and Sun \cite{HsiehS08} based their work on Aha et al.'s CBR model \cite{LTW} and used StarCraft replays to construct states and building sequences (``build orders''). Schadd et al. \cite{SchaddBS07} applied a CBR approach to opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS game (a ``Total Annihilation'' clone). Finally, \v{C}ertick\'{y} et al. \cite{certicky2013cbr} used CBR to build their army, based on the opponent's army composition, and they pointed out on the importance of proper scouting for better results.

@@

% MACHINE LEARNING:
Concerning tactical decision making, many different approaches have been explored such as machine learning or game tree search. 
Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters in first person shooter games (FPS) units tracking. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. \cite{SynnaeveTactics} combined space abstraction into regions from \cite{Perkins10} and tactical-decision making by assigning scores (economical, defenses, etc.) to regions and looking for their correspondences to tactical moves (attacks) in pro-gamers replays.
{\color{blue} Finally, Christopher Miles \cite{miles2006co} created the idea of {\em IMTrees}, a tree where each leaf node is an influence map, and each intermediate node is a combination operation (sum, multiplication); Miles used evolutionary algorithms to learn IMTrees for each strategic decision in the game involving spatial reasoning by combining a set of basic influence maps.}

% GAME TREE SEARCH:
Game tree search techniques have also been explored for tactical decision making. \cite{Chung05} applied Monte-Carlo planning to a capture-the-flag version of Open RTS. Balla and Fern \cite{UCT} applied the UCT algorithm (a Monte Carlo Tree Search algorithm) to tactical assault planning in Wargus. To make game tree search applicable at this level, abstract game state representations are used in order to reduce the complexity. Also, abstractions, or simplifications about the set of possible actions to execute in a given game state need to be used.

@@

Holistic approaches to address RTS AI attempt to address the whole problem using a single unified method. To the best of our knowledge, with a few exceptions, such as the Darmok system \cite{OntanonMSR10} (which uses a combination of case-based reasoning and learning from demonstration) or ALisp \cite{Marthi05}, there hasn't been much work in this direction.   The main reason is that the complexity of RTS games is too large, and approaches that decompose the problem into smaller, separate, problems, achieve better results in practice. However, holistic approaches, based, for example, on Monte Carlo Tree Search, have not been explored so far, and would be a very interesting venue of future work.

>>>>

Another way  to measure the complexity of the game is  by looking at the branching factor, $b$, and the depth of the game, $d$, as proposed in~\cite{Gaby}, with a total game complexity of $b^d$. In Chess, $b \approx 35$ and $d \approx 80$. In more complex games, like Go, $b \approx 30 - 300$, and $d \approx 150 - 200$. In order to determine the branching factor in StarCraft when an AI plays it, we must have in mind, that the AI can issue actions simultaneously to as many units in the game as desired. Thus, considering that, in a typical game, a player controls between 50 to 200 units, the branching factor would be between $u^{50}$ and $u^{200}$, where $u$ is the average number of actions each unit can execute. 
{\color{blue}
Estimating the value of $u$ is not easy, since the number of actions a unit can execute is highly dependent on the context. Let us make the following assumptions: 1) at most 16 enemy units will be in range of a friendly unit (larger values are possible, but are unlikely), 2) when an AI plays StarCraft, it only makes sense to consider movement in the 8 cardinal directions per unit (instead of assuming that the player can issue a ``move'' command to anywhere in the map at any point in time), and 3) for ``build'' actions, we consider that SCVs only build in their current location (otherwise, if they need to move, we consider that as first issuing a ``move'' action, and then a ``build''). With those assumptions, units in StarCraft can execute between 1 (units like ``Supply Depots'', whose only action is to be ``idle'') to 43 actions (Terran ``Ghosts''), with typical values around 20 to 30. Now, if we have in mind that actions have cool-down times, and thus not all units can execute all of the actions at every frame, we can take a conservative estimation of about $10$ possible actions per unit per game frame. This results in a conservative estimate for the branching factor between $b \in [10^{50},10^{200}]$, only considering units (ignoring the actions buildings can execute). 
}
%Assuming a conservative value for $u$ of about $30$ (moving to nearby positions, attacking units in range, using special abilities, etc.), we obtain a branching factor of $b$ between  $30^{50}$ and $30^{200}$.
Now, to compute $d$, we simply consider the fact that typical games last for about 25 minutes, which results in $d \approx 36000$ (25 minutes $\times$ 60 seconds $\times$ 24 iterations per second).

@@

% CBR:
Also falling into the machine-learning category, a significant group of researchers has explored case-based reasoning (CBR) \cite{Aamodt94CBR} approaches for strategic decision making. For example Aha et al. \cite{LTW} used CBR to perform dynamic plan retrieval in the Wargus domain. Hsieh and Sun \cite{HsiehS08} based their work on Aha et al.'s CBR model \cite{LTW} and used StarCraft replays to construct states and building sequences (``build orders''). Schadd et al. \cite{SchaddBS07} applied a CBR approach to opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS game (a ``Total Annihilation'' clone). Jaidee et al. \cite{jaidee2011case} study the use of CBR for automatic goal selection, while playing an RTS game. These goals will then determine which Q-tables to be used in a reinforcement learning framework. Finally, \v{C}ertick\'{y} et al. \cite{certicky2013cbr} used CBR to build their army, based on the opponent's army composition, and they pointed out on the importance of proper scouting for better results.

@@

% MACHINE LEARNING:
Concerning tactical decision making, many different approaches have been explored such as machine learning or game tree search. 
{\color{blue} Hladky and Bulitko \cite{Hladky2008} benchmarked hidden semi-Markov models (HSMM) and particle filters for unit tracking, although the used first-person shooter (FPS) games for their experimentation, the results apply to RTS games as well}. They showed that the accuracy of occupancy maps was improved using movement models (learned from the player behavior) in HSMM. Kabanza et al. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN, used for plan and intent recognition to find tactical opportunities. Sharma et al. \cite{CBR-RL} combined CBR and reinforcement learning to enable reuse of tactical plan components. Cadena and Garrido \cite{CadenaG11} used fuzzy CBR (fuzzy case matching) for strategic and tactical planning. \cite{SynnaeveTactics} combined space abstraction into regions from \cite{Perkins10} and tactical-decision making by assigning scores (economical, defenses, etc.) to regions and looking for their correspondences to tactical moves (attacks) in pro-gamers replays.
{\color{blue} Finally, Christopher Miles \cite{miles2006co} created the idea of {\em IMTrees}, a tree where each leaf node is an influence map, and each intermediate node is a combination operation (sum, multiplication); Miles used evolutionary algorithms to learn IMTrees for each strategic decision in the game involving spatial reasoning by combining a set of basic influence maps.}

% GAME TREE SEARCH:
Game tree search techniques have also been explored for tactical decision making. Churchill and Buro \cite{churchill2012fast} presented the ABCD algorithm (Alpha-Beta Considering Durations), a game tree search algorithm for tactical battles in RTS games. Chung et al. \cite{Chung05} applied Monte-Carlo planning to a capture-the-flag version of Open RTS. Balla and Fern \cite{UCT} applied the UCT algorithm (a Monte Carlo Tree Search algorithm) to tactical assault planning in Wargus. To make game tree search applicable at this level, abstract game state representations are used in order to reduce the complexity. Also, abstractions, or simplifications about the set of possible actions to execute in a given game state need to be used.

@@

Holistic approaches to address RTS AI attempt to address the whole problem using a single unified method. To the best of our knowledge, with a few exceptions, such as the Darmok system \cite{OntanonMSR10} (which uses a combination of case-based reasoning and learning from demonstration) or ALisp \cite{Marthi05}, there hasn't been much work in this direction.   The main reason is that the complexity of RTS games is too large, and approaches that decompose the problem into smaller, separate, problems, achieve better results in practice. However, holistic approaches, based, for example, on Monte Carlo Tree Search, have only been explored in the context of smaller-scale RTS games~\cite{ontanon2013naive}, but techniques that scale up to large RTS games like StarCraft are still not known.
